---
title: "Exam 1"
author: "Sam"
date: "2022-10-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1

```{r}
df_1 = read.table("webtraffic.txt", header = TRUE)
```

#### 1(A)

```{r}
df_1 = read.table("webtraffic.txt", header = TRUE)
# Create count matrix
df_1 <- colSums(df_1)
Traffic <- matrix(data = 0, nrow = 9, ncol = 9)
Traffic[9, 1] <- 1000
k = 1
for (i in 1:8) {
  for (j in 1:9) {
    Traffic[i, j] <- df_1[k]
    k = k + 1
  }
}

# Create probability matrix
Traffic_probability <- t(apply(Traffic ,1 , function(x) x/sum(x)))
Traffic
```

#### 1(B)

```{r}
set.seed(6)
library(markovchain)

states <- c("1", "2", "3", "4", "5", "6", "7", "8", "9")

mc <- new("markovchain", states=states, transitionMatrix=Traffic_probability)

dimnames(Traffic_probability) <- list(states, states)
mc@transitionMatrix <- Traffic_probability

par(mar=c(0.1 ,0.1 , 0.1, 0.1))

plot(mc, edge.arrow.size=0.5, 
     vertex.size = 25, 
     edge.label.cex = 0.0001)
```

#### 1(C)

```{r}
Traffic_probability
```

#### 1(D)

The probability of a visitor being on Page 5 after 5 clicks is `r Traffic_probability[5, 5]`

#### 1(E)

```{r}
Traffic_probability[9, 1] = 1; Traffic_probability[9, 9] = 0
Q=t(Traffic_probability) - diag(9)
Q[9,] = rep(1, 9)
rhs = c(rep(0, 8), 1)
Pi = solve(Q,rhs)
print(Pi)
```

#### 1(F)

```{r}
B=Traffic_probability[1:8,1:8]
Q=diag(8)-B
rhs=c(0.1, 2, 3, 5, 5, 3, 3, 2)
m=solve(Q,rhs)
```

The average time a visitor spends on the website until he/she first leaves is `r m[1]` minutes.

# Question 2

#### 2(A)

$$
n \ge \frac{var[p(x)]}{(tolerance)^2\delta} = \frac{\frac{1}{\lambda^2}}{(10^{-3})^2*0.01}
\\
n \ge \frac{10^8}{\lambda^2}
$$

#### 2(B)

#### Lambda = 1

```{r}
lambda = 1

n = 10^8/lambda^2
x = runif(n, 0, 1)
y = -log(x)/lambda
g = sin(y)/lambda
result = sum(g)/n
result_real = 1 / (1+lambda^2)
print(result)
print(result_real)
```

Using MCMC, we get a value of `r result`. This is different from the real result, `r result_real` by `r abs(result-result_real)`. Therefore it is within the tolerance.

#### Lambda = 2

```{r}
lambda = 2

n = 10^8/lambda^2
x = runif(n, 0, 1)
y = -log(x)/lambda
g = sin(y)/lambda
result = sum(g)/n
result_real = 1 / (1+lambda^2)
print(result)
print(result_real)
```

Using MCMC, we get a value of `r result`. This is different from the real result, `r result_real` by `r abs(result-result_real)`. Therefore it is within the tolerance.

#### Lambda = 4

```{r}
lambda = 4

n = 10^8/lambda^2
x = runif(n, 0, 1)
y = -log(x)/lambda
g = sin(y)/lambda
result = sum(g)/n
result_real = 1 / (1+lambda^2)
print(result)
print(result_real)
```

Using MCMC, we get a value of `r result`. This is different from the real result, `r result_real` by `r abs(result-result_real)`. Therefore it is within the tolerance.