---
title: "HW 02"
author: "Group 10"
date: "2022-10-03"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Homework 2

##### Read csv
```{r,echo = TRUE}
auto = read.csv("Auto.csv", na.strings = "NA")
```

##### 1(a)
```{r}
auto$origin = factor(auto$origin, 1:3, c("US", "Europe", "Japan"))

freq <- table(auto$origin)
barplot(freq)
```

##### 1(b)
```{r}
lmb <- lm(mpg ~ origin + weight + year, data = auto)
par(mfrow=c(2,2))
plot(lmb)
```


The plots indicate that:
- The error term is not normally distributed
- There is variance is not constant, thus heteroskedasticity is present

##### 1(c)
```{r}
lmc <- lm(log(mpg) ~ origin + log(weight) + year + I(year^2), data = auto)
par(mfrow=c(2,2))
plot(lmc)
```

```{r}
summary(lmc)
```

The model assumptions seem to have been roughly satisfied now.

##### 1(d)
```{r}
plot(auto$year, log(auto$mpg), 
     main="Log(mpg) vs year", 
     xlab = "Year", ylab="MPG")

min = -coef(lmc)[5] / (2 * coef(lmc)[6])
```

The relationship appears U-shaped based on the plot above. The minimum is `r min`.

##### 1(e)
```{r}
summary(lmc)
par(mfrow = c(1, 2))
plot(auto$weight, auto$mpg)
plot(log(auto$weight), log(auto$mpg))
```

It tells us that as you increase the weight the mpg falls. The relationship for the unlogged version is similar, less linear, but still negative.

##### 2(a)
$$
y_i = \gamma_0 + \gamma_1(x_i-\bar{x}) + \gamma_2(x_i-\bar{x})^2 + e_i = 
\\
\gamma_0 + y_1x_i - y_1\bar{x} + y_2x_i^2 - 2y_2x_i\bar{x} + \gamma_2\bar{x}^2 + e_i = 
\\
(\gamma_0 - \gamma_1\bar{x} + \gamma_2\bar{x}^2) + (\gamma_1 - 2\gamma_2\bar{x})x_i + \gamma_2x_i^2 + e_i 
\\
\therefore \beta_0 = \gamma_0 - \gamma_1\bar{x} + \gamma_2\bar{x}^2
\\
\beta_1 = \gamma_1 - 2\gamma_2\bar{x}
\\
\beta_2 = \gamma_2
$$

##### 2(b)
```{r}
auto$year_squared <- auto$year^2
summary(lm(mpg ~ year + year_squared, data = auto))
```

##### 2(c)
The correlation between year and year squared is `r cor(auto$year, (auto$year)^2)`.

##### 2(d)
The mean of year is `r mean(auto$year)`.

##### 2(e)
```{r}
auto$year_centered <- auto$year - mean(auto$year)
auto$year_centered_squared <- (auto$year_centered)^2
```

The correlation between centered year and centered year squared is `r cor(auto$year_centered, (auto$year_centered)^2)`

##### 2(f)
```{r}
par(mfrow = c(1, 2))
plot(auto$year, auto$year_squared, main = "(Year vs Year Squared)", xlab = "Year", ylab = "Year Squared")
plot(auto$year_centered, auto$year_centered_squared, main = "Centered (Year vs Year Squared)", xlab = "Centered Year", ylab = "Centered Year Squared")
```

##### 2(g)
```{r}
lm_2g <- lm(mpg ~ year_centered + year_centered_squared, data = auto)
summary(lm_2g)
```

##### 2(h)

$$
\beta_0 = \gamma_0 - \gamma_1\bar{x} + \gamma_2\bar{x}^2
\\
\beta_1 = \gamma_1 - 2\gamma_2\bar{x}
\\
\beta_2 = \gamma_2
$$
```{r}
gamma_0 <- coef(lm_2g)[1]
gamma_1 <- coef(lm_2g)[2]
gamma_2 <- coef(lm_2g)[3]
mean_year <- mean(auto$year)
```

<center>$\beta_0 =$ `r gamma_0 - gamma_1 * mean_year + gamma_2 * mean_year^2`<br /></center>
<center>$\beta_1 =$ `r gamma_1 - 2 * gamma_2 * mean_year`<br /></center>
<center>$\beta_2 =$ `r gamma_2`<br /></center>