---
title: "airlineEDA_markdown"
output:
  pdf_document: default
  html_document:
    df_print: paged
date: "2022-08-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)   
library(ggplot2) 
setwd("C:/Users/a1184/MSiA_Director_Volunteer")
```

## Read in Datasets

Read in dataset January2019 and July2019 (provided)

```{r}
January2019 <- read.csv("On_Time_Reporting_Carrier_On_Time_Performance_2019_1.csv",
                        header = TRUE, sep = ",")
July2019 <- read.csv("On_Time_Reporting_Carrier_On_Time_Performance_2019_7.csv",
                     header = TRUE, sep = ",")
```

See column names:
```{r}
names(January2019)
```




## Data Selection

columns to keep - total 29

```{r}
keep_names <- c("Year", "Month", "DayofMonth", "DayOfWeek", "DepTime", "CRSDepTime",
                "ArrTime", "CRSArrTime", "Reporting_Airline","Flight_Number_Reporting_Airline", 
                "Tail_Number", "ActualElapsedTime", "CRSElapsedTime", "AirTime", "ArrDelay", "DepDelay", "Origin",
                "Dest","Distance", "TaxiIn", "TaxiOut", "Cancelled",
                "CancellationCode", "Diverted", "CarrierDelay", "WeatherDelay", "NASDelay", "SecurityDelay",
                "LateAircraftDelay")

#Note: DayOfWeek - 1 is Monday
```

Select columns, and randomly sample 100K observations from each dataframe
```{r}
temp <- select(January2019, all_of(keep_names))
names(temp)
# rename some of the variables
temp <- temp %>% rename(UniqueCarrier = Reporting_Airline,
                        FlightNum = Flight_Number_Reporting_Airline,
                        TailNum = Tail_Number)
temp100k <- sample_n(temp, 100000)
jan <- temp100k
```

```{r}
temp <- select(July2019, all_of(keep_names))
names(temp)
temp <- temp %>% rename(UniqueCarrier = Reporting_Airline,
                        FlightNum = Flight_Number_Reporting_Airline,
                        TailNum = Tail_Number)
temp100k <- sample_n(temp, 100000)
jul <- temp100k

rm(temp); rm(temp100k)
```
## End Data Selection

## EDA

```{r}
head(jan, n=5)
head(jul, n=5)
```

Compute missing data proportion

```{r}
sapply(jan, function(col)
  sum(is.na(col)) / length(col)
) %>% round(digits=2)
sapply(jul, function(col)
  sum(is.na(col)) / length(col)
) %>% round(digits=2)

# What do you observe
```

```{r}
select(jan, ArrDelay, CarrierDelay:LateAircraftDelay) %>% head(n=10)
select(jul, ArrDelay, CarrierDelay:LateAircraftDelay) %>% head(n=20)
```


For a comprehensive view on handling missing data in R see here:
https://rmisstastic.netlify.app/

## Flight Data

Number of flights

```{r}
par(mfrow=c(2,2))
barplot(table(jan$DayofMonth), main="January: Day")
barplot(table(jan$DayOfWeek), main="January: Weekday (1=Monday)")
barplot(table(jul$DayofMonth), main="July: Day")
barplot(table(jul$DayOfWeek), main="July: Weekday (1=Monday)")
```

Departure Time

```{r}
par(mfrow=c(1,2))
hist(jan$CRSDepTime, main = "January: Scheduled Departure Time")
hist(jul$CRSDepTime, main = "July: Scheduled Departure Time")
```

```{r}
par(mfrow=c(2,3))
hist(jan$DepDelay, main = "January: Departure Delay, minutes")
hist(jan$ArrDelay, main = "January: Arrival Delay, minutes")
plot(jan$DepDelay, jan$ArrDelay, main = "January")
abline(0,1)

hist(jul$DepDelay, main =  "July: Departure Delay, minutes")
hist(jul$ArrDelay, main = "July: Arrival Delay, minutes")
plot(jul$DepDelay, jul$ArrDelay)
abline(0,1)
```

```{r}
par(mfrow=c(1,1))
plot(jan$CRSDepTime, jan$ArrDelay, 
     main = "January: Arrival Delay vs Scheduled Departure Time")

```

```{r}
head(jan$CRSDepTime)
```

Distance and AirTime

```{r}
par(mfrow=c(1,2))
plot(jan$Distance, jan$AirTime)
jan[ which(jan$AirTime>500) , ]

plot(jul$Distance, jul$AirTime)
jul[ which(jul$AirTime>500) , ]
```

## Feature Engineering

Weekday

```{r}
jan$weekday <- factor(jan$DayOfWeek, 
                      labels=c("Mon","Tue","Wed","Thu","Fri","Sat","Sun"))
jul$weekday <- factor(jul$DayOfWeek, 
                      labels=c("Mon","Tue","Wed","Thu","Fri","Sat","Sun"))
```

log-transform delay times

```{r}
summary(jan$ArrDelay)
```

```{r}
#pmax function returns the parallel maxima of two or more input vectors.
jan$ArrDelay <- pmax(0, jan$ArrDelay)  
jan$DepDelay <- pmax(0, jan$DepDelay)
jul$ArrDelay <- pmax(0, jul$ArrDelay)
jul$DepDelay <- pmax(0, jul$DepDelay)
summary(jan$ArrDelay)
#Address zeros for logs
jan$logArrDelay <- log(1 + jan$ArrDelay)
jan$logDepDelay <- log(1 + jan$DepDelay)
jul$logArrDelay <- log(1 + jul$ArrDelay)
jul$logDepDelay <- log(1 + jul$DepDelay)
summary(jan$logArrDelay)
```

Transform Departure Time

```{r}
# %% is the modulo operation
jan$DecimalDepTime <- floor(jan$CRSDepTime / 100) + 
  (jan$CRSDepTime %% 100)/60
head(jan$CRSDepTime)
head(jan$DecimalDepTime)
```

```{r}
jul$DecimalDepTime <- floor(jul$CRSDepTime / 100) + 
  (jul$CRSDepTime %% 100)/60
head(jul$CRSDepTime)
head(jul$DecimalDepTime)
```

Average Speed

```{r}
par(mfrow=c(1,2))
jan$speed <- jan$Distance / (jan$AirTime/60)
boxplot(jan$speed)
jul$speed <- jul$Distance / (jul$AirTime/60)
boxplot(jul$speed)
```

```{r}
filter(jan, speed >= 650)
#remove speeds greater than 650 mph
jan_low_speed <- filter(jan, speed <= 650)
```

Flight delay by time of day

```{r}
#Add a smoothed conditional mean geom_smooth
ggplot(jan, aes(x=DecimalDepTime, y=ArrDelay)) +
  geom_point() + geom_smooth()
ggplot(jan, aes(x=DecimalDepTime, y=logArrDelay)) +
  geom_point(alpha=0.8, shape = ".") + geom_smooth(colour = "darkgreen")
#alpha is the transparency
#geom_smooth uses gam for more than 1000 points
```

```{r}
ggplot(jul, aes(x=DecimalDepTime, y=ArrDelay)) +
  geom_point() + geom_smooth()
ggplot(jul, aes(x=DecimalDepTime, y=logArrDelay)) +
  geom_point(alpha=0.8, shape = ".") + geom_smooth(colour = "darkgreen")
```
















